Contracts for Contracts

### 2023-06-26 paper/case_study.tex
intro
more formal stuff

examining 3-4 cases, they are the proof that the approach works

@@

>>>>

We now turn our attention to real world cases and examine how \lang
can be used to consolidate smart contract programs.
These case studies are empirical evidence showing \lang is effective
in preventing attacks.

@@

\paragraph{Severity} talk about severity for each case

### 2023-06-26 paper/formal_model.tex
more formal stuff

\section{Formal Model} \label{sec:model}

@@

\Cref{fig:syntax} shows the abstract syntax of \corelang,
which models the most important portion of Solidity.
\corelang's type contains integers, unsigned integers, booleans,
addresses, mappings, and struct.
A mapping type has the domain type and codomain type.
A struct type has a name and a list of type-and-identifier declarations.

At the top-level, a contract $c$ consists of a list of type-and-identifier
declarations and a list of function definitions.  As in other languages, a
function definition has a name, parameter declarations, return types, and the
function body.  To model boundaries between multiple smart contracts, a
function can be either \Keywd{public} or \Keywd{private}.
In \corelang, a specification $\sigma$ is additionally attached to every function.

@@

Expressions could be an identifier, a literal value (e.g. numbers), a binary
operation, projections of mappings or structs, address calls, and function calls.
An address call has form $e_1.f\Keywd{\{}(x: e_2)^*\Keywd{\}(}e_3^*\Keywd{)}$,
where $e_1$ is the target expression of type address,
$(x: e_2)^*$ are call options (e.g. specifying the gas budget),
and $e_3^*$ are call arguments.
An ordinary function call or struct constructor call is simpler, consisting
only of the function name and argument expressions.

Specifications attached to functions first introduce bindings for the guarded
function, including the function name, option arguments, ordinary arguments,
and returned values.
The precondition and postcondition are denoted in the \code{requires}-clause and
\code{ensures}-clause, respectively.
In the \code{where}-clause, programmers can specify the condition for addresses
appeared as arguments or returned values for the parent call, recursively using
the same syntax.
In practice not all functions need to be specified with conditions.
In such cases, a no-op specification is expressed where the precondition and
postcondition are simply a boolean expression \code{true}.

@@

The syntax permits arbitrary expressions appeared as preconditions and
postconditions.
However, not all possible expression are valid executable specifications.
For example, a programmer can write \code{requires x + 1} as a precondition,
which do have a meaning for our purpose.
The static semantics presented here concerns about when a specification is
considered valid.

@@

@@

@@

Our translation semantics identifies the boundary

>>>>

\section{Formal Semantics} \label{sec:model}

@@

\Cref{fig:syntax} shows the abstract syntax of \corelang,
which models the most important portion of Solidity.
\corelang's type universe contains integers, unsigned integers, booleans,
addresses, mappings, and struct.
A mapping type has the domain type and codomain type.
A struct type has a name and a list of type-and-identifier declarations.

At the top-level, a contract $c$ consists of a list of type-and-identifier
declarations and a list of function definitions.
Those type-and-identifier declarations specify storage states.
As in other languages, a function definition has a name, parameter
declarations, return types, and the function body.  To model boundaries between
multiple smart contracts, a function can be either \Keywd{public} or
\Keywd{private}.
In \corelang, a specification $\sigma$ is additionally attached to every function.

@@

Expressions could be an identifier, a literal value (e.g. numbers), a binary
operation, projections of mappings or structs, address/function calls.
Address and function calls have form $\AddrCall{e_1}{(x: e_2)^*}{e_3}$,
where $e_1$ is the target callee expression,
$(x: e_2)^*$ are call options (e.g. specifying the gas budget),
and $e_3^*$ are call arguments.
An ordinary function call or struct constructor call omits the call option
part, consisting only of the function name and argument expressions.

Specifications attached to function definitions first introduce bindings for
the guarded function, including the function name, option parameters,
ordinary parameters, and returned values.
The precondition and postcondition are denoted in the \code{requires}-clause and
\code{ensures}-clause, respectively.
In practice not all functions need to be specified with conditions.
In such cases, a no-op specification is expressed where the precondition and
postcondition are simply a boolean expression \code{true}.
In the \code{where}-clause, programmers can specify the condition for addresses
appeared as arguments or returned values for the parent call, recursively using
the same syntax.
When the specification is attached to an address, we allow very flexible ways
to specify the callee. For example, if a struct argument contains an address
value that we would like to guard:
\begin{lstlisting}
// x is a struct argument whose field y is an address
... where { x.y.f requires $\mathit{precond}$ ensures $\mathit{postcond}$ }
\end{lstlisting}
we can use the projection syntax to specify
the condition when address \code{x.y} is called on function \code{f}.

@@

The syntax permits arbitrary expressions appeared as preconditions and
postconditions.
However, not all possible expression are valid executable specifications.
For example, a programmer can write arithmetic expressions \code{requires x + 1} as a precondition,
which do have a meaning for our purpose.
The static semantics presented here concerns about when a specification is
considered valid.

@@

\begin{figure}[t]
  \judgement{Function}{$\Gamma \vdash \sigma: t$}
  %\begin{alignat*}{2}
    %\mathit{param}(\SpecCall{f}{(x : y)^*}{x^*}{y^*} ...) & = 1
  %\end{alignat*}

  \infrule[f-spec]{
    \mathit{target}(\sigma) = f \\
    \Gamma' = \Gamma \cup
              \{ \mathit{param}(\sigma)_i : \mathit{type}(d_i) \}^{i \in 1..m} \\
    \Gamma'' = \Gamma' \cup
              \{ \mathit{ret}(\sigma)_j : t_j \}^{j \in 1..n} \\
    \Gamma' \vdash \mathit{pre}(\sigma) : \Keywd{bool} \andalso
    \Gamma'' \vdash \mathit{post}(\sigma) : \Keywd{bool} \\
    \Gamma \vdash \mathit{addr}(\sigma)_i : \note{type-of-call}
  }{
    \Gamma \vdash \sigma : \FunType{f}{d_i^{i \in 1..m}}{m}{t_j^{j \in 1..n}}
  }

  \infrule[c-top]{
    \forall i \in 1..n,\ \Gamma_0 \cup \Gamma \vdash \F_i \\
    \Gamma = \{ \mathit{id}(d_j) : \mathit{type}(d_j) \}^{j \in 1..m} \cup
             \{ \mathit{id}(\F_i) : \mathit{type}(\F_i) \}^{i \in 1..n}
  }{
    \Gamma_0 \vdash \Contract{c}{d_i^{i \in 1..m}}{{\F}_i^{i \in 1..n}}
  }

  % \Gamma_0 initial context, containing typings for built-in terms

  \caption{Static semantics (excerpt) of \corelang specifications. Only checkings relevant to specifications are shown.}
  \label{fig:static}
\end{figure}

@@

\note{GW: what if in the condition, user calls a guarded function?}

\paragraph{Extent of Effective Checking}
monitors all value flows within the extent

\paragraph{Wrapping Addresses} \lang transform \Keywd{address} type to a struct
type with an additional type $\mathcal{G}$ that carries the pre-/postcondition.
At runtime when the underlying address is called,
corresponding checks can be performed by examining this value of type $\mathcal{G}$.

$$
\Keywd{address} \leadsto \Keywd{address} \times \mathcal{G}
$$

\paragraph{Unwrapping Addresses} balabala

@@

Our translation semantics identifies the extent of effective checking.
Notably, if an address value is stored to the state or passed to other
contracts, the address leaves the boundary and loses its guards by unwrapping.

\corelang's semantics guarantee relatively complete monitoring, i.e.
